%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Thesis Template @ The University of Manchester
% LaTeX Template
% Version 1 (24/08/2018)
% Joe Crone
%
% This template is based on:
% The University of Manchester, Presentation of Thesis Policy
% Research Office Graduate Education Team
% June 2017
% http://www.regulations.manchester.ac.uk/pgr-presentation-theses/
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%



% Document Class
\documentclass[11pt,oneside]{thesisformat}

% Packages
\usepackage{subfiles}

% Thesis Information
\title{Thesis Template, The University of Manchester}
\def\thesistitle{Design and Optimisation of High-Energy Inverse Compton Scattering Sources Driven by Multi-Pass Energy Recovery Linacs}
\def\name{Joe Crone}
\def\faculty{Faculty of Science and Engineering}
\def\school{School of Natural Sciences}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

% Title Page
%--------------------------------------------------------
\maketitle


% Contents Page (+ word count)
%--------------------------------------------------------
\tableofcontents
\vfill
\begin{center}
\quickwordcount{main}
\end{center}
\vfill

% List of Tables
%--------------------------------------------------------
\listoftables


% List of Figures
%--------------------------------------------------------
\listoffigures


% List of Abbreviations
%--------------------------------------------------------

% [optional] Lay Abstract
%--------------------------------------------------------
%\chapter*{Lay Abstract}
%\addcontentsline{toc}{chapter}{Lay Abstract}


% Declaration
%--------------------------------------------------------
\chapter*{Declaration}
\addcontentsline{toc}{chapter}{Declaration}

No portion of the work referred to in the thesis has been submitted in support of an application for another degree or qualification of this or any other university or other institute of learning.


% Copyright Statement
%--------------------------------------------------------
\chapter*{Copyright Statement}
\addcontentsline{toc}{chapter}{Copyright Statement}

\begin{enumerate}[(i)]
\item The author of this thesis (including any appendices and/or schedules to this thesis) owns certain copyright or related rights in it (the  “Copyright”) and s/he has given The University of Manchester certain  rights to use such Copyright, including for administrative purposes.
\item Copies of this thesis, either in full or in extracts and whether in hard or electronic copy, may be made only in accordance with the Copyright, Designs and Patents Act 1988 (as  amended) and regulations issued under it or, where appropriate, in accordance with licensing agreements which the University has from time to time. This page must form part of any such copies made.
\item The ownership of certain Copyright, patents, designs, trademarks and other intellectual property (the “Intellectual Property”) and any reproductions of copyright works in the thesis, for example graphs and tables (“Reproductions”), which may be described in this thesis, may not be owned by the author and may be owned by third parties. Such Intellectual Property and Reproductions cannot and must not be made available for use without the prior written permission of the owner(s) of the relevant Intellectual Property and/or Reproductions.
\item Further information on the conditions under which disclosure, publication and commercialisation of this thesis, the Copyright and any Intellectual Property  and/or Reproductions described in it may take place is available in the University IP Policy (see \url{http://documents.manchester.ac.uk/DocuInfo.aspx?DocID=24420}), in any relevant Thesis restriction declarations deposited in the University Library, The University Library’s regulations (see \url{http://www.library.manchester.ac.uk/about/regulations/}) and in The University’s policy on Presentation of Theses.
\end{enumerate}

% Abstract
%--------------------------------------------------------
\chapter*{Abstract}
\addcontentsline{toc}{chapter}{Abstract}

% TOO LONG!

% Little bit about ICS sources - why offer narrowband radiation, why offer high photon energy, why limited in flux, why use electrons? 

High quality x-ray sources are required for fundamental physics research into areas such as atomic physics and material science. Third generation Synchrotron light source facilities fulfill this need in most aspects; producing a high flux (number of photons) at x-ray energies up to 100's~\si{\kilo\electronvolts} with high brightness and a narrow bandwidth (energy spread of the produced photons). However, the maximum photon energy produced by synchrotrons is limited by facility size, electron beam energy and magnet strength constraints. High energy x-rays are required for experimentation with thick specimens and can be advantageous for some atomic physics experiments. % statement about synchrotron access + local experiments? 
An x-ray inverse Compton scattering source allows for production of high-flux narrowband x-rays on the lab-scale with fluxes and bandwidths between conventional bremmsstrahlung x-ray sources and synchrotron light sources. However, inverse Compton sources offer the advantageous access to higher energy x-rays than those available at third generation synchrotron radiation sources. 

Similarly, high quality $\gamma$-ray sources are in demand for experimentation in nuclear photonics, nuclear forensics and proliferation and for medical isotope production. Plans for the proposed flagship ELI-NP-GBS $\gamma$-ray source are indicative of such demand. Currently inverse Compton scattering sources, such as HI$\gamma$S, produce $\gamma$-rays with world leading photon fluxes and narrow bandwidths. Alternatively, bremmstrahlung $\gamma$-ray sources could be used though these are inherently broadband and monochromation of $\gamma$-rays is poor. Radioactive isotope sources could also be used though these have two primary disadvantages: they are typically low flux and can't be turned off. Therefore, ICS production is the premier method of $\gamma$-ray production for research yet higher flux and more narrowband sources are required.

Electron beams used to drive ICS sources can be provided by a variety of electron accelerators, though high fluxes of photons are only obtainable with a high average electron beam current (high repetition rate and bunch charge) and small emittance whilst narrow bandwidths require small emittance and small electron beam energy spreads. Currently, most ICS sources utilise either storage rings, which provide high average beam current and moderate electron bunch energy spread, or linacs, with small emittance and energy spread. Energy recovery linacs (ERLs) -- where linac quality electron bunches are re-circulated --  could be ideal because these could provide electron beams with small emittance, energy spread and high average beam current simultaneously. Hence, within this thesis ICS sources driven by ERLs are investigated for production of high energy photons beyond those available at synchrotron radiation sources.      

In this thesis two inverse Compton scattering sources are designed: the CBETA x-ray ICS source and the DIANA $\gamma$-ray ICS source. % what should I say here?
Methods for predicting the flux and the produced photon spectrum are developed and a series of optimisations toward maximal narrowband photon production are proposed. 
% needs expanding, not sure how


% [optional] Dedication / Acknowledgement
%--------------------------------------------------------
\chapter*{Acknowledgement}
\addcontentsline{toc}{chapter}{Acknowledgement}



% [optional] Preface / 'The Author' - for the external examiner
%--------------------------------------------------------
%\chapter*{Preface}
%\addcontentsline{toc}{chapter}{Preface}


% Introduction
%--------------------------------------------------------
\subfile{Chapters/Introduction}

% Part 1:
%--------------------------------------------------------
% Theory of Energy Recovery Linac Design, \subfile only likes single word .tex names
\subfile{Chapters/Energy_Recovery_Linac_Design}

% Part 2:
%--------------------------------------------------------
% Photon Production by Inverse Compton Scattering
\subfile{Chapters/Photon_Production_by_Inverse_Compton_Scattering}


% Part 3:
%--------------------------------------------------------
% Optimisation and Charaterisation of Inverse Compton Scattering Spectra
\subfile{Chapters/Optimisation_and_Characterisation_of_Inverse_Compton_Scattering_Spectra}


% Part 3:
%--------------------------------------------------------
% CBETA Multipass Commissioning
%\subfile{Chapters/CBETA_Multi-Pass_Commissioning}

% Part 4:
%--------------------------------------------------------
% CBETA ICS Design
\subfile{Chapters/CBETA_Inverse_Compton_Source_Design}

% Part 5:
%--------------------------------------------------------
% DIANA ICS Design
\subfile{Chapters/DIANA_Inverse_Compton_Source_Design}



% Conclusion
%--------------------------------------------------------
\subfile{Chapters/Conclusion}

% Bibliography
%--------------------------------------------------------
%\addcontentsline{toc}{chapter}{Bibliography}
\printbibliography
\bibliographystyle{unsrt}
\bibliography{sample}

% Appendix
%--------------------------------------------------------
\appendix

\include{Appendices/AppendixTemplate}


\end{document}
